---
title: "SEC Data analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose of this project {#Purpose of this project}

### SEC Data analysis with R

The primary purpose of this project is to retrieve company data from SEC filings and analyze it for educational and research purposes. Please note that the scripts provided here are not intended for making investment decisions, and their use is at your own risk.

## DISCLAIMER {#sec-disclaimer}

*This financial data retrieval code is provided for informational purposes only, and we make no warranties regarding the accuracy, completeness, or timeliness of the data. It should not be considered financial advice, and users should consult with qualified professionals for personalized guidance. Data is obtained from various sources, including the SEC, and we do not guarantee its accuracy or availability. Users are responsible for assessing and managing risks associated with the financial data. We are not liable for any damages arising from the use of this data or code. Redistribution, resale, or republication of the data without authorization is prohibited. This disclaimer is subject to change, and users are responsible for staying updated. It is not legal advice, and compliance with applicable laws and regulations is the user's responsibility.*

## Setup libraries {#sec-setup-libraries}

Import required libraries (install if necessary):

```{r Libraries}
library(tidyverse)
library(httr)
library(jsonlite)
```

Sourcing required files in your main_script.R (here commented)

```{r Files}
source("../Functions/data_retrieval.R")
source("../Functions/data_analysis.R")
source("../Functions/data_visualization.R")
source("../Functions/utils.R")
```

## Retrieve data from SEC {#sec-retrieve-data-from-sec}

We start by defining the user agent for accessing the SEC API. Next, we retrieve the full list of companies, and then select the first company in the list `$cik[1]` to retrieve its data. The cik for this example is Apple Inc. "AAPL"

```{r Retrieve data from SEC}
# Define user headers
headers <- c('User-Agent' = 'email@address.com')

# Retrieve company list
company_List <- retrieve_Company_List(headers)

# Retrieve company information
company_List[1,] %>% print()
cik <- company_List$cik_str[1]
company_Data <- retrieve_Company_Data(headers, cik)

```

## Explore data retrieved from SEC {#sec-explore-data-retrieved-from-sec}

We split the data in `company_Data` into separate Lists: `company_Metadata`, `company_Facts`, `company_Concept`

```{r Explore Data}
company_Metadata <- company_Data$company_Metadata
company_Facts <- company_Data$company_Facts
company_Concept <- company_Data$company_Concept
```

Let's create now a first dataframe:

```{r Explore Data_2}
# Create a DataFrame with relevant data
company_df <- data.frame(
  CIK = cik,
  Name = company_Metadata$name,
  FiscalYearEnd = company_Metadata$fiscalYearEnd,
  AssetsLabel = company_Facts$facts$`us-gaap`$Assets$label,
  AssetsDescription = company_Facts$facts$`us-gaap`$Assets$description
  # Add more relevant fields here
)
print(company_df)
```

### Structure of the SEC data: XBRL

The structure the data retrieved from SEC Api is a structured financial data in compliance with SEC reporting standards. It's organized this way to ensure consistency and comparability of financial data across different companies and filings.

In practice, if you want to analyze specific financial metrics like "Assets," you would navigate through the structure to access the corresponding "fact" under the relevant "concept." This allows you to extract and work with financial data for analysis or reporting.In the structure you provided:

-   **`cik`**: This is the Central Index Key (CIK) of the company. It's a unique identifier assigned to each company that files reports with the U.S. Securities and Exchange Commission (SEC).

-   **`entityName`**: This is the name of the company, in this case, "Apple Inc."

-   **`facts`**: This is the main container for financial data. Within "facts," you have two main sections:

    -   **`dei`**: This section contains financial data related to Document and Entity Information (DEI). DEI data includes basic company information, such as the entity's common stock shares outstanding and the entity's public float.

    -   **`us-gaap`**: This section contains financial data that follows the U.S. Generally Accepted Accounting Principles (GAAP). It includes a wide range of financial concepts, which are organized as a list of financial items.

Within these sections, you'll find the actual financial data represented as a list of items. Each item has specific attributes that describe the financial information:

-   **Concept**: A "concept" represents a specific financial measure or item. For example, "Assets" or "AccountsPayable" are concepts. Each concept has an associated label and a data type.

-   **Facts**: Under each concept, you have "facts." These are the actual numerical values or data associated with the concept. For example, "Assets" will have a fact that represents the total assets value for the company. Facts typically have attributes like "value," "unitRef," and "contextRef."

-   **Attributes**: Facts may have additional attributes that provide context for the data, such as the reporting period, currency unit, or the precision of the data.

To analyze and work with this data effectively, you would typically need to select specific concepts and facts that are relevant to your analysis, and you may also need to transform or pivot the data into a more tabular format for further processing and visualization.

### Company Metadata

General information about the company is included in the Metadata. This includes its name and fiscal year end date and other information.

```{r Metadata}
company_Metadata <- company_Data$company_Metadata
company_Metadata$name        # Company name
company_Metadata$fiscalYearEnd  # Fiscal year end date
names(company_Metadata) %>% print() # Full list of Metadata
```

### Company Facts

Financial facts reported by the company are included in Facts. Access to them is based on their label, e.g. "Assets".

```{r Facts}
company_Facts <- company_Data$company_Facts

assets_fact <- company_Facts$facts$`us-gaap`$Assets # Example of data from us-gaap section

str(assets_fact) # it shows the structure of the List "asset_fact"
```

This list contains information related to "Assets" and appears to be organized as follows:

1.  **label**: The description of the item, which in this case is "Assets." This label provides a brief identifier for the financial concept.

2.  **description**: A more detailed description of the item, explaining what "Assets" represent. In this case, it mentions that it is the sum of carrying amounts as of the balance sheet date of all recognized assets. It also describes assets as probable future economic benefits obtained or controlled by an entity due to past transactions or events.

3.  **units**: This section contains information about the units in which the data is presented. In this case, it appears to be in USD (U.S. Dollars).

4.  **USD**: Within the "units" section, there is a nested data frame with several columns. Each row in this data frame corresponds to a specific piece of information, potentially for different fiscal quarters.

    -   **end**: The date corresponding to the end of the fiscal quarter for which the data is reported. For example, "2008-09-27" represents the end of the quarter.

    -   **val**: The value of "Assets" for that specific fiscal quarter. It's given in numeric format, e.g., "3.96e+10," which is a scientific notation for a large number.

    -   **accn**: A unique identifier associated with this financial data, often in the format of "Accession Number." This number is used to reference specific financial reports.

    -   **fy**: The fiscal year associated with the data. For instance, "2009" is the fiscal year for some of the reported values.

    -   **fp**: Indicates the fiscal period (e.g., "Q3" for the third quarter and "FY" for the full fiscal year).

    -   **form**: The form type associated with the financial data report (e.g., "10-Q" for a quarterly report and "10-K" for an annual report).

    -   **filed**: The date when the financial report was filed, often in the format "YYYY-MM-DD."

    -   **frame**: An additional identifier, potentially associated with the fiscal quarter or financial reporting period. For example, "CY2008Q3I" could be related to the calendar year 2008, the third quarter, and might have an "I" indicating it's the initial or amended report.

This structure contains detailed information about "Assets" reported by a company over multiple fiscal quarters, including dates, values, report identifiers, and more. You can use this data to analyze the company's financial performance over time or create visualizations to understand how its assets have evolved.

```{r Facts_2}
assets_label <- assets_fact$label          # Label of the fact ("Assets")
assets_description <- assets_fact$description  # Description of the fact
assets_unit <- assets_fact$units

print("Label:")
print(assets_fact$label)

print("Description:")
print(assets_fact$description)

```

```{r Facts_3}

print("Units (USD):")
print(assets_fact$units$USD)

```

The following are the list of the Facts in the us-gaap section:

```{r Fact_us-gaap_list}
# Get the names of elements within us-gaap and indicate if the list is truncated
names_us_gaap <- names(company_Facts$facts$`us-gaap`)
truncated <- length(names_us_gaap) > 20
if (truncated) {
  names_us_gaap <- head(names_us_gaap, 20)
  cat("The list is truncated.\n")
}

# Print the names and the number of elements
print(names_us_gaap)
cat("The number of Facts in us-gaap section is:", length(company_Facts$facts$`us-gaap`))

```

The following are the list of the Facts in the deo section:

```{r Fact_dei_list}
# Get the names of elements within dei
names_dei <- names(company_Facts$facts$dei)

# Print the names and the number of elements
print(names_dei)
cat("The number of Facts in dei section is:", length(names_dei))

# Print an example element
cat("Entity Common Stock Shares Outstanding is:")
```

```{r Facts_dei_EntityCommonStockSharesOutstanding}

print(company_Facts$facts$dei$EntityCommonStockSharesOutstanding$units$shares)
```

```{r Facts_dei_list_2}

# Print an example element
cat("Entity Public Float (USD) is:")
```

```{r Facts_dei_list_EntityPublicFloat}
print(company_Facts$facts$dei$EntityPublicFloat$units$USD)
```
